Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-14 paper/current/main.tex
update abstract

Human mobility and migration drives major social phenomenon such as the growth and evolution of cities~\autocite{bettencourt2007cities, bettencourt2013cities, wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity, hufnagel2004epidemics, eubank2004outbreaks}, economies~\autocite{kaluza2010cargo, kerr2011immigration, taylor2016refugees, pappalardo2015mobility}, innovation~\autocite{braunerhjelm2020labor, kaiser2018innovation, armano2017innovation, liu2010spillovers},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, jonkers2013return, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity, simini2012universal}, describes mobility using the \textit{distance} between places.
However, amid a globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization}, geographic distance has less relevance for understanding mobility, whereas other abstract notions of distance~\autocite{boschma2005proximity} pose practical and theoretical challenges.
Using the case of scientific mobility, we leverage the neural embedding model~\autocite{mikolov2013word2vec} to demonstrate a data-driven approach for learning a vector-space embedding that encodes a functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} between organizations and captures the latent and multi-faceted structure of human mobility.
We train neural embeddings of scientific organizations using mobility trajectories derived from the papers published between 2008 and 2019 of disambiguated authors from the Web of Science database.
Functional distance is derived as the cosine distance between vectors in the resulting embedding and is shown to better explain and predict real mobility between organizations than geographic distance. 
Exploiting their unique properties, we interpret the embedding space and observe how it captures latent factors related to culture, language, and prestige of organizations at multiple levels of granularity.
This approach can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data, and can easily be generalized to other mobility data.
\end{abstract}

>>>>

Human mobility and migration drives major social phenomenon such as the growth and evolution of cities~\autocite{wef2017migration, curiel2018citygravity},  epidemics~\autocite{kraemer2020covid, truscott2012epidemicgravity, xia2005measlesgravity}, economies~\autocite{kaluza2010cargo, kerr2011immigration}, innovation~\autocite{kaiser2018innovation, armano2017innovation},  and scientific research~\autocite{sugimoto2017mostimpact, petersen2018multiscale, franzoni2014advantage, morgan2018prestige, rodrigues2016mobility}. 
Common models of mobility, such as the Gravity Model~\autocite{zipf1946gravity, xia2005measlesgravity, jung2008highwaygravity, barbosa2018human, curiel2018citygravity, lewer2008immigrationgravity, hong2016busgravity, truscott2012epidemicgravity, xia2005measlesgravity}, describes mobility using the \textit{distance} between places.
However, geographic distance is less relevent in the increasingly-globalized globalized world~\autocite{czaika2018migration, czaika2018globalisation, buchan2009globalization}, in which mobility is dependent less on geography than linguistic, economic, cultural, and historical similarity. 
Alternative and abstract notions of distance attempt to capture the latent landscape of mobility~\autocite{boschma2005proximity,  brown1970functional, brown1970migration, kim2018functional} but pose practical and theoretical challenges. 
Here, using the case of scientific mobility, we leverage the neural embedding model~\autocite{mikolov2013word2vec} to demonstrate a data-driven approach for learning a vector-space embedding that encodes a functional distance~\autocite{brown1970migration, brown1970functional, kim2018functional} between organizations and captures the latent and multi-faceted structure of scientist's mobility.
We demonstrate that the cosine distance between organization vectors in the embedding, constructed using the affiliation trajectories of scientists in the Web of Science database, better explain and predict real mobility than geographic distance. 
Exploiting their unique properties, we interpret the embedding space and observe how it captures latent factors related to culture, language, and prestige of organizations at multiple levels of granularity.
This approach can learn scalable, dense, and meaningful vector representations of scientific mobility directly from the data, and can easily be generalized to other mobility data.
\end{abstract}
